# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r loaddata}
unzip("activity.zip")
activity <- read.csv("activity.csv", colClasses = c("integer", "Date", "character"))
na_steps <- sum(is.na(activity$steps))
na_date <- sum(is.na(activity$date))
na_interval <- sum(is.na(activity$interval))
summary(activity)
```

We can see that there are `r na_steps` data points with NA for the number of steps taken.   Therefore the next step
is to remove the NA data from the data set.

```{r cleandata}
activity_clean <- activity[!is.na(activity$steps), ]
summary(activity_clean)
```


## What is mean total number of steps taken per day?


```{r tableofmeans,results="asis"}
steps_by_day <- tapply(activity_clean$steps, activity_clean$date, sum)
library(xtable)
print(xtable(as.data.frame(steps_by_day), caption="Daily Total number of steps"), type="html")
```

Here is a histogram:

```{r histogram}
hist(steps_by_day, main="Histogram showing the spread of the total number of steps per day", breaks = c(0, 500, 5000, 10000, 15000, 20000, 25000), freq=TRUE)
mean_steps <- mean(steps_by_day)
median_steps <- median(steps_by_day)
```

We have an average of `r mean_steps` steps being taken per day, with a median of `r median_steps`

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
