# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

```{r loaddata}
unzip("activity.zip")
activity <- read.csv("activity.csv", colClasses = c("integer", "Date", "integer"))
na_steps <- sum(is.na(activity$steps))
na_date <- sum(is.na(activity$date))
na_interval <- sum(is.na(activity$interval))
summary(activity)
```

We can see that there are `r na_steps` data points with NA for the number of steps taken.   Therefore the next step
is to remove the NA data from the data set.   As we may need the original set later on, we create a seperate data set with the cleaned up data.

```{r cleandata}
activity_clean <- activity[!is.na(activity$steps), ]
summary(activity_clean)
```

## What is mean total number of steps taken per day?

Firstly we need to calculate the total number of steps taken per day, and then we need to get the mean and median of that data set.
```{r tableofmeans,results="asis"}
steps_by_day <- tapply(activity_clean$steps, activity_clean$date, sum)
library(xtable)
## print(xtable(as.data.frame(steps_by_day), caption="Daily Total number of steps"), type="html")
```

We're asked to produce a histogram:

```{r histogram,echo=FALSE,warnings=FALSE}
options(scipen = 1, digits = 0)
hist(steps_by_day, main="Histogram showing the spread of the total number of steps per day", )
mean_steps <- mean(steps_by_day)
median_steps <- median(steps_by_day)
```

We have an average of `r mean_steps` steps being taken per day, with a median of `r median_steps`

## What is the average daily activity pattern?

The following plot shows the average number of steps taken in each interval, averaged over all the days.

```{r intervalmeans, echo=FALSE}
steps_by_interval <- tapply(activity_clean$steps, activity_clean$interval, mean)
plot(rownames(steps_by_interval), steps_by_interval, type = "l")
```

The time interval with the maximum average number of steps is `r names(steps_by_interval[steps_by_interval == max(steps_by_interval)])`

## Inputing missing values

The assigned scheme is to use the average time for the interval to replace the NA values.   As mentioned above there are `r na_steps` data points with NA for the number of steps taken.  

```{r fileinnas}
filledin_data <- activity
for (key in rownames(steps_by_interval)) {
  filledin_data[is.na(filledin_data$steps & filledin_data$interval == key), ]$steps <- steps_by_interval[key]
}
filled_steps_by_day <- tapply(filledin_data$steps, filledin_data$date, sum)
filled_mean_steps <- mean(filled_steps_by_day)
filled_median_steps <- median(filled_steps_by_day)
hist(filled_steps_by_day, main="Total number of steps per day with filled in values", )
```

The revised mean is `r filled_mean_steps` and the revised median is
`r filled_median_steps`

## Are there differences in activity patterns between weekdays and weekends?

For this section, I assume that we use the filled in data set for this section.
```{r weekday}

# We create a new column in the data set, daytype.   Default is "weekday", but then
# rewrite as "weekend" where the day is either Sunday or Saturday

filledin_data$daytype <- "weekday"
filledin_data$daytype[weekdays(filledin_data$date) == "Sunday"] <- "weekend"
filledin_data$daytype[weekdays(filledin_data$date) == "Saturday"] <- "weekend"

weekday_steps_by_interval <- tapply(filledin_data[filledin_data$daytype=="weekday",]$steps, filledin_data[filledin_data$daytype=="weekday",]$interval, mean)
weekend_steps_by_interval <- tapply(filledin_data[filledin_data$daytype=="weekend",]$steps, filledin_data[filledin_data$daytype=="weekend",]$interval, mean)

par(mfrow=c(2,1))
plot(rownames(weekend_steps_by_interval), weekend_steps_by_interval, type="l", ylab="Steps in interval", xlab="Interval", main="Weekend")
plot(rownames(weekday_steps_by_interval), weekday_steps_by_interval, type="l", ylab="Steps in interval", xlab="Interval", main="Weekday")

# for the panel plot we need two sets of means - one for the weekday and one
# for the weekend.

```
